### 2017-01-21 thesis/methods/methods.tex
Finished obesity metagene prediction section

\section{Obesity metagene prediction}
\label{sec:obesity_metagene_prediction}

% TODO: possibly extra analyses here (gradual pathway inclusion in the model, using all the pathway...?)
Sample BMI, BMI status and/or pathway metagenes were used to construct a linear model in Cris Print's data.
The constructed linear model was then used to predict the obesity metagene values in \citet{Creighton2012} data set.
The predicted obesity metagene values were compared with the true obesity metagene values, and both Spearman and Pearson correlation were calculated.

@@

% TODO: re-write everything so that it's in past tense

>>>>

\section{Obesity metagene prediction with pathway metagenes}
\label{sec:obesity_metagene_prediction_with_pathway_metagenes}

To determine whether the obesity metagene was associated with any of the pathway metagenes, a linear model was constructed with some clinical variables and pathway metagenes, and then this model was used to predict the obesity metagene.
This was based on the idea that, if a variable or a pathway metagene was a significant part of the model for the obesity metagene, then the variable or a pathway metagene may be associated with the obesity metagene.
Furthermore, the model that was created based on the significant variables can be used to predict the obesity metagene, which can be used to compare with the original obesity metagene; if the predicted metagene was similar to the original metagene, then it provides an evidence that the pathway used to construct the model was associated with the obesity metagene.

Four linear models were created in (Cris' data -- get reference) to predict the obesity associated metagene from the \citet{Creighton2012} study.
The first model used the sample \gls{bmi} value or \gls{bmi} status by itself, the second model with sample \gls{bmi} value, \gls{bmi} status and selected pathway associated metagenes, the third model with selected pathway associated metagenes, and the last model was constructed with just the \gls{pr} pathway.
The selected pathways used in the linear model construction were \gls{bcat}, \gls{er}, \gls{ifna}, \gls{ifny}, Myc and \gls{pr}.

In deciding on the pathways to be used for the construction of the linear model, the correlation of different versions of pathway metagenes, \gls{svd}-derived and transformation matrix-derived, in different data sets were used.
When a transformation matrix is applied to the data set in which it was derived from, the resulting metagene is identical to the one created from the direct application of \gls{svd}.
Therefore, theoretically, if the metagene derived from a transformation matrix is similar to the metagene created from \gls{svd}, then it shows that the genetic signature is consistent and reliable across different data sets.
The above mentioned pathways had high correlation value between the \gls{svd}-derived and transformation matrix-derived pathway metagenes, and so it was used in the construction of the linear model.

% TODO: possibly extra analyses here (gradual pathway inclusion in the model, using all the pathway...?)

The predicted obesity metagenes that were created in (Cris' data -- get reference) was compared with the original obesity metagene from the \citet{Creighton2012} data set.
Both Pearson and Spearman correlation was calculated to assess the relationship between the predicted and original metagenes.
Scatter plots were also created to visually compare the two metagene scores.

@@
